### 2018-08-28 paper/preliminaries.tex
re-worked all definitions and proofs

\begin{example}\label{example:CSP}
  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$.

@@

In fact, in every semantic model, subset containment is used to define refinement. The
model we focus on first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
events in which $P$ may refuse to engage, after having performed the events of
$s$. The failures model of a process $P$ records all its failures in a set
$\fails(P)$. 
Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
the set 
$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the  refusals
of $P$ after the
trace $s$. 
Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.

@@

For finite CSP processes, since the refusals of each process state are
subset-closed, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}; these are defined by 
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals $\refs(P/s)$ by  
%
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
The cardinality of $\maxrefs(P/s)$ reflects the degree of
nondeterminism that is present in process state $P/s$: the more maximal refusal sets
contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply 
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}

@@

The main idea of the underlying test strategy for failures refinement can be
based on solving a \emph{hitting set problem}. Given a finite collection of
finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
satisfying the following property.
%
\begin{equation}
  \label{eq:hit}
  \forall A\in C: H\cap A \neq\varnothing.
\end{equation}
%
A \emph{minimal hitting set} is a hitting set that cannot be further reduced
without losing the characteristic property (\ref{eq:hit}). The problem of
determining minimal hitting sets is known to be
NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces the
effort of testing for failures refinement from a factor of $2^\Sigma$ to a
factor that equals the number of minimal hitting sets.

For testing, the following lemma about hitting sets are required.

\begin{lemma}
\label{lemma:hseta}
Let $P, Q$ be two finite-state CSP processes.
%% satisfying $P\lessdet_T Q$.
For each $s\in\trc(P)$,
let $\text{minHit}(P/s)$ denote the
collection of all minimal hitting sets of $\minaccs(P/s)$.
Then the following statements are equivalent.
\begin{enumerate}
\item $P\ conf\ Q$

\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
\end{enumerate}
\end{lemma}
\begin{proof}
For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and  
  $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), states that
\[
\forall A_Q\in\minaccs(G(Q)/s):
\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
\]
Therefore, $H \in  \text{minHit}(P/s)$ not only implies $H\cap
A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq
A_Q$ for at least one $A_P$. As a consequence, each $H \in
\text{minHit}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.

To prove ``$2 \Rightarrow 1$'', assume that that Statement~2 of the lemma holds but 
that  $P\ conf\ Q$ does {\it not} hold. 
According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. 
Minimising $\overline H$ yields   a minimal hitting set $H\in
\text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%

>>>>

\begin{example}\label{example:CSP}
  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$. By convention, these choices evaluate to the process $\Stop$ 
which does not engage into any event, if the index set $I$ is empty.

@@

In fact, in every semantic model, subset containment is used to define refinement. The
model we focus on first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
events in which $P$ may refuse to engage, after having performed the events of
$s$. The failures model of a process $P$ records all its failures in a set
$\fails(P)$. The semantic rules specify for each operator, how the failures of 
the resulting process can be calculated from the failures of each operand. For example, 
$\fails(P\intchoice Q) = \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for the comprehensive list of these rules.

Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
the set 
$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the  refusals
of $P$ after the
trace $s$. 
Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.

@@

For finite CSP processes, since the refusals of each process state are
subset-closed, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}; these are defined by 
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals $\refs(P/s)$ by  
%
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.
The maximal refusals in combination with the initials of a process express its degree of 
nondeterminism.
\begin{example}
\label{ex:nondetdegree}
Process $P = (\Stop \intchoice Q)$ has maximal refusals 
$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ 
refuses to engage into any event, and this carries over to $P$, when applying the internal 
choice operator. However, $P$ is distinguished from $\Stop$ by its initials: 
$[P]^0 = [ \Stop \intchoice Q]^0 = [Q]^0$, so $P$ may engage nondeterministically in
any initial event of $Q$, but can also refuse everything, due to internal selection of the $\Stop$ process.

Assuming that $\Sigma = \{a,b,c,d\}$, process
$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$ has maximal refusals
$\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials $[Q]^0=\Sigma$, so, in contrast to
$P$, nondeterminism is reflected
here by the presence of two maximal refusals.
\xbox
\end{example}

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
{\it possibly} refusing the events from $[P/s]^0 - A$ and 
{\it always} refusing those from $\Sigma - [P/s]^0$.  
The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply 
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}

@@

The main idea of the underlying test strategy for failures refinement can be
based on solving a \emph{hitting set problem}. Given a finite collection of
finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
satisfying the following property.
%
\begin{equation}
  \label{eq:hit}
  \forall A\in C: H\cap A \neq\varnothing.
\end{equation}
%
A \emph{minimal hitting set} is a hitting set that cannot be further reduced
without losing the characteristic property (\ref{eq:hit}). By $\minhits(C)$ we denote the collection of minimal hitting sets for $C$. For the pathological case where $C$ contains an empty set, $\minhits(C)$ is also empty.
 The problem of
determining minimal hitting sets is known to be %%% \cite{Book1975-BOOKRM,5533149}
NP-hard~\cite{5533149}, but we will see below that it reduces the
effort of testing for failures refinement from a factor of $2^{|\Sigma|}$ to a
factor that equals the number of minimal hitting sets.

The following lemma establishes the fact that the $conf$ relation specified in (\ref{eq:conf}) can be chracterised by means of minimal acceptances and their
minimal hitting sets.
\begin{lemma}
\label{lemma:hseta}
Let $P, Q$ be two finite-state CSP processes.
%% satisfying $P\lessdet_T Q$.
For each $s\in\trc(P)$,
let $\minhits(P/s)$ denote the
collection of all minimal hitting sets of $\minaccs(P/s)$.
Then the following statements are equivalent.
\begin{enumerate}
\item $P\ conf\ Q$

\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
\end{enumerate}
\end{lemma}
\begin{proof}
For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and  
  $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), states that
\[
\forall A_Q\in\minaccs(G(Q)/s):
\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
\]
Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap
A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq
A_Q$ for at least one $A_P$. As a consequence, each $H \in
\minhits(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.

To prove ``$2 \Rightarrow 1$'', assume that that Statement~2 of the lemma holds but 
that  $P\ conf\ Q$ does {\it not} hold. 
According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. 
Minimising $\overline H$ yields   a minimal hitting set $H\in
\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%

It is important to point out that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$. Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we conclude that 
$\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{ \varnothing \}$. These considerations
clarify that $\minhits(P/s)$ is empty if and only if  $\minaccs(P) = \{ \varnothing \}$.
We note that the proof of Lemma~\ref{lemma:hseta} also covers the situations where
$\minaccs(P/s) = \{ \varnothing \}$ and therefore  $\minhits(P/s) = \varnothing$.
